---
title: "Introduction to R, Session 2 <html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>"
author: "Rockefeller University, Bioinformatics Resource Centre"
date: "http://rockefelleruniversity.github.io/Intro_To_Python/"
output: 
  xaringan::moon_reader:
    css: ["default", "metropolisCustom.css", "metropolis-fontsCustom.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
  html_document:
    toc: true # table of content true
    toc_float: yes
    depth: 3  # upto three depths of headings (specified by #, ## and ###)
    number_sections: false  ## if you want number sections at each table header
    theme: united  # many options for theme, this one is my favorite.
    highlight: tango  # specifies the syntax highlighting style
params:
  isSlides: "no"
---

```{r,include=FALSE}
suppressPackageStartupMessages(require(knitr))
knitr::opts_chunk$set(echo = TRUE, tidy = T)
```

```{r, results='asis',include=TRUE,echo=FALSE}
if(params$isSlides != "yes"){
  cat("# Introduction to Python

---
"    
  )
  
}

```

---
##  Recap

?? can add a recap to link to keypoints in first session??
Session 1 covered introduction to R data types and import/export of data.

- [Background to R](https://rockefelleruniversity.github.io/Intro_To_R_1Day/r_course/presentations/singlepage/introToR_Session1.html#background-to-r)
- [Data types in R](https://rockefelleruniversity.github.io/Intro_To_R_1Day/r_course/presentations/singlepage/introToR_Session1.html#data_types_in_r)
- [Reading and writing in R](https://rockefelleruniversity.github.io/Intro_To_R_1Day/r_course/presentations/singlepage/introToR_Session1.html#reading-and-writing-data-in-r)


---
## Control Statements

There are several ways to control how your code is evaluated. There are two main classes:

Conditional branching (if,else)
Loops (for, while)


<span style="color:green">While</span> I'm analysing data, <span style="color:red">if</span> I need to execute complex statistical procedures on the data I will use Python <span style="color:red">else</span> I will use a calculator.

---
## Conditional Branching

Conditional branching is the evaluation of a logical to determine whether a chunk of code is executed.

In Python, we use the if statement with the logical to be evaluated immediately after. The  dependent code is indicated by a colon, new line and indentation.

```{python}

x = True
if x:
  print("x is true")

```

```{python}

x = False
if x:
  print("x is true")

```

---
## Conditional Branching

More often, we construct the logical value within the if statement itself. This can be termed the condition.

```{python}

x = 10
y = 4
if x > y:
  print("The value of x is",x,"which is greater than", y)

```
The message is printed above because x is greater than y.


```{python}
y = 20
if x > y:
  print("The value of x is",x,"which is greater than", y)

```

x is now no longer greater than y, so no message is printed.

We really still want a message telling us what was the result of the condition.

---
## else following an if

If we want to perform an operation when the condition is false we can follow the *if* statement with an *else* statement.


```{python}

x = 3
if x < 5:
  print(x, "is less than to 5")
else:
  print(x, "is greater than or equal to 5")

```

```{python}

x = 10
if x < 5:
  print(x, "is less than to 5")
else:
  print(x, "is greater than or equal to 5")

```

---
## else if = elif

We may wish to execute different procedures under multiple conditions. This can be controlled using the *elif* following an initial *if* statement.

```{python}

x = 5
if x < 5:
  print(x, "is less than to 5")
elif x > 5:
  print(x, "is greater than 5")
else:
  print(x, "is 5")

```

---
## Loops

While and for loops iterate over a block of code, and keep rerunning it. 

*While* loops do this while a specific condition is met (or until that condition is not met).

*For* loops will do this for a given number of iterations.

---
## While Loop

While loops have a similar structure to if statements. We start by designating the while loop, ten follow with the logical to be evaluated immediately after. The  dependent code is indicated by a colon, new line and indentation.

```{python}

x = 1
while x < 3:
  print("x is",x)
  x = x+1

```

---
## For Loop

For loops do not have a condtional. Instead you supply an object that you want to be iterate over. This can be a list, tuple, dictionary, set or string. Here we use a list. 

```{python}

x = ['Alpha','Bravo','Charlie']
for i in x:
  print(i)

```


---
## for and range()

The range() function provides us with a nice input for our for loops. It returns a sequence of numbers, starting from 0 and stops before the specified number.

```{python}
for i in range(3):
  print("i is", i)

```

---
## Looping through indices

When we have a numeric range, we can use it to index out from existing objects. This often allows for more complex code evaluation. 

```{python}

geneName = ["Ikzf1","Myc","Igll1"]
expression = [10.4, 4.3, 6.5]
iterations = len(geneName)

for i in range(iterations):
  print(geneName[i]," has an TPM of ",expression[i])

```


---
## Loops and conditionals

Loops can be combined with conditional statements to allow for complex control of their execution over Python objects. Though these can be tough to read and create, by keeping an eye of the hierarchy of identation we can follow the logic. 

```{python}
x = range(8)

for i in range(8):
  if i > 5:
    print("Number",i,"is greater than 5")
  elif i == 5:
    print("Number",i,"is  5") 
  else:
    print("Number",i,"is less than 5") 

```

---
## Breaking loops

We can use conditionals to exit a loop if a condition is satisfied, just like a while loop.

```{python}
x = range(8)

for i in range(8):
  if i > 5:
    print("Number",i,"is greater than 5")
  elif i == 5:
    print("Number",i,"is  5")
    break
  else:
    print("Number",i,"is less than 5") 

```



