---
title: "Introduction to Python, Session 2 <html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>"
author: "Rockefeller University, Bioinformatics Resource Centre"
date: "http://rockefelleruniversity.github.io/Intro_To_Python/"
output: 
  xaringan::moon_reader:
    css: ["default", "metropolisCustom.css", "metropolis-fontsCustom.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
  html_document:
    toc: true # table of content true
    toc_float: yes
    depth: 3  # upto three depths of headings (specified by #, ## and ###)
    number_sections: false  ## if you want number sections at each table header
    theme: united  # many options for theme, this one is my favorite.
    highlight: tango  # specifies the syntax highlighting style
params:
  isSlides: "no"
---

```{r,include=FALSE}
suppressPackageStartupMessages(require(knitr))
knitr::opts_chunk$set(echo = TRUE, tidy = T)
```

```{r, results='asis',include=TRUE,echo=FALSE}
if(params$isSlides != "yes"){
  cat("# Introduction to Python

---
"    
  )
  
}

```

```{r, eval=T, echo=F, warning=F, message=F, hide=T}
library(IntroToPython)
library(reticulate)
py_install(c("matplotlib", "numpy", "scipy","seaborn"), pip_options = "--quiet")
```

---
##  Recap

?? can add a recap to link to keypoints in first session??
Session 1 covered introduction to R data types and import/export of data.

- [Background to Python](https://rockefelleruniversity.github.io/Intro_To_Python/presentations/singlepage/Session1.html#Background_of_Python)
- [Set up](https://rockefelleruniversity.github.io/Intro_To_Python/presentations/singlepage/Session1.html#Set_Up)
- [Variables and Functions](https://rockefelleruniversity.github.io/Intro_To_Python/presentations/singlepage/Session1.html#Variables_and_Functions)
- [Data Objects](https://rockefelleruniversity.github.io/Intro_To_Python/presentations/singlepage/Session1.html#Data_Objects)
- [Custom Functions](https://rockefelleruniversity.github.io/Intro_To_Python/presentations/singlepage/Session1.html#Custom_Functions)

---
## Session Overview

- [Course Home Page](http://rockefelleruniversity.github.io/Intro_To_Python/)
- [Reading and Writing Files](https://rockefelleruniversity.github.io/Intro_To_Python/presentations/singlepage/Session2.html#Reading_and_Writing_Files)
- [Control Statements](https://rockefelleruniversity.github.io/Intro_To_Python/presentations/singlepage/Session2.html#Control_Statements)

---
```{r, results='asis',include=TRUE,echo=FALSE}
if(params$isSlides == "yes"){
  cat("class: inverse, center, middle

# Reading and Writing Files

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=720px></html> 

---
"    
  )
}else{
  cat("# Reading and Writing Files
      
---
"    
  )
  
}

```


## Reading/writing files

Most of the time, you will not be generating data in Python but will be importing data from external files.

A standard format for this data is a table:


```{r,echo=F}
minRep <- rbind(cbind(matrix(rnorm(12,4),ncol=3,byrow = T),matrix(c(rnorm(9,4),rnorm(3,8)),ncol=3,byrow = T)),
cbind(matrix(rnorm(12,10),ncol=3,byrow = T),matrix(c(rnorm(6,3),rnorm(6,10)),ncol=3,byrow = T)))
colnames(minRep) <- paste0(c("Sample_"),
                      1:5,".",sort(rep(c("hi","low"),3)))
minRepdf <- data.frame(Gene_Name=paste("Gene",letters[1:8],sep="_"),minRep)
#minRepdf$Gene_Name <- paste("Gene",letters[1:8],sep="_")
#write.table(minRepdf,file="readThisTable.csv",sep=",",row.names=F)
kable(minRepdf[,1:4], format='html')

```

---
## Working Directory

If you are reading in data, you need to know from which vantage point your computer is looking. The view point of your computer is the Working Directory. 

There are several utilities in the os (operating system) package. First lets check our current Working Directory using `getcwd`. 


```{python}
import os
os.getcwd()

```

---
## Working Directory

To set our Working Directory we can use the `chdir` function. Let's navigate to the course material that we downloaded. There are some example data sets there. 

```{python, eval=F}

os.chdir("/Users/mattpaul/Downloads/Intro_To_Python-master/r_course/")

```

```{python, eval=T}

os.listdir()

```

```{python, eval=T}

os.listdir("data")

```


---
## Reading

There are many ways to read in data. Most of the time we are importing simple 2D tables, so we want to generate a NumPy array using the `genfromtxt()` function.

Here we are reading a csv file (Comma-Separated Values). This means each value in our file is separated by a comma. So when we read the file we will specify that the delimiter is a comma. 


```{python}
import numpy as np
my_table = np.genfromtxt("data/ToRead.csv", delimiter=",")

type(my_table)
```

```{python}
my_table
```

---
## Reading

Here we try to read in a more complex file. This data is a mixture of characters and numbers. The `dtype` argument can be used to specify the [format of the data](https://numpy.org/doc/stable/reference/arrays.dtypes.html#arrays-dtypes) we are reading in. This data also has column titles, so we will skip over those with `skip_header`. 

The data set contains: 

* **Sex** - **U**nicode string
* **Height** - **f**loat
* **Weight** - **f**loat

```{python}

dtype = ['U10' , 'f', 'f' ]

my_table2 = np.genfromtxt('data/height-weight.csv', delimiter = ",", skip_header=True, dtype=dtype)

```

---
## Reading

When you read in an array with multiple data formats it does not automatically make a 2D array, instead a 1D array containing other arrays. 

```{python}
my_table2

```

---
## Reading 

While reading we can also assign our data into multiple objects using the `unpack` argument. Here we assign each column into it's own separate array. 

```{python}

sex, height, weight = np.genfromtxt('data/height-weight.csv', unpack = True, delimiter = ",", skip_header=True, dtype=dtype)

sex

```


---
## Reading 

An alternative to setting the data type explicitly, you can let genfromtxt determine it automatically by setting the dtype to `None`. This works pretty well most of the time.  

```{python}

sex, height, weight = np.genfromtxt('data/height-weight.csv', delimiter = ",", skip_header=True, unpack = True, dtype=None)


```


---
## Quick Recap

Once you have performed some analysis you will want to export it into a new file. 

Let's do some quick processing based on what we learned yesterday. First we will make a 2D array. 

```{python}

new_array = np.array([sex, height, weight])
```

Next we do a logical test to see which entries in uor first row are Female.

```{python}
to_subset = new_array[0,:]=="Female"
to_subset
```

---
## Quick Recap

We can use the boolean array we created to subset our array, based on whether they are `Female` or not. 


```{python}

subset_array = new_array[:,to_subset]
subset_array 

```

Then export the subset array with the `savetxt` function. This has similar arguments to `genfromtxt`. A key addition is `fmt`, the format in which the data is saved. You can specify scientific notation, significant digits or in our case simply that we will use a string. 

```{python}

np.savetxt("height-weight_female.csv", subset_array, delimiter=',', fmt='%s')

```

---
## Complex reading/writing

Though a lot of the time we may use these NumPy approaches, there are specific functions dedicated to a variety of data types. 

The most common is [pandas](https://pandas.pydata.org/docs/getting_started/index.html#getting-started). This is a specialized library for managing data frames. It also has the ability to read/write from excel spreadsheets.  

[BioPython](https://biopython.org/) has a range of utilities for managing biological data types i.e. fasta, fastq, etc.



---
```{r, results='asis',include=TRUE,echo=FALSE}
if(params$isSlides == "yes"){
  cat("class: inverse, center, middle

# Control Statements

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=720px></html> 

---
"    
  )
}else{
  cat("# Control Statements
      
---
"    
  )
  
}

```


---
## Control Statements

There are several ways to control how your code is evaluated. There are two main classes:

* Conditional branching (if,else)  
* Loops (for, while)


<span style="color:green">While</span> I'm analyzing data, <span style="color:red">if</span> I need to execute complex statistical procedures on the data I will use Python <span style="color:red">else</span> I will use a calculator.

---
## Conditional Branching

Conditional branching is the evaluation of a logical to determine whether a chunk of code is executed.

In Python, we use the if statement with the logical to be evaluated immediately after. Thedependent code is indicated by a colon, new line and indentation.

```{python}

x = True
if x:
  print("x is true")

```

```{python}

x = False
if x:
  print("x is true")

```

---
## Conditional Branching

More often, we construct the logical value within the if statement itself. This can be termed the condition.

```{python}

x = 10
y = 4
if x > y:
  print("The value of x is",x,"which is greater than", y)

```
The message is printed above because x is greater than y.


```{python}
y = 20
if x > y:
  print("The value of x is",x,"which is greater than", y)

```

x is now no longer greater than y, so no message is printed.

We really still want a message telling us what was the result of the condition.

---
## else following an if

If we want to perform an operation when the condition is false we can follow the `if` statement with an `else` statement.


```{python}

x = 3
if x < 5:
  print(x, "is less than to 5")
else:
  print(x, "is greater than or equal to 5")

```

```{python}

x = 10
if x < 5:
  print(x, "is less than to 5")
else:
  print(x, "is greater than or equal to 5")

```

---
## else if = elif

We may wish to execute different procedures under multiple conditions. This can be controlled using the `elif` following an initial `if` statement.

```{python}

x = 5
if x < 5:
  print(x, "is less than to 5")
elif x > 5:
  print(x, "is greater than 5")
else:
  print(x, "is 5")

```

---
## Loops

While and for loops iterate over a block of code, and keep rerunning it. 

`While` loops do this while a specific condition is met (or until that condition is not met).

`For` loops will do this for a given number of iterations.

---
## While loop

While loops have a similar structure to if statements. We start by designating the while loop, ten follow with the logical to be evaluated immediately after. The  dependent code is indicated by a colon, new line and indentation.

```{python}

x = 1
while x < 3:
  print("x is",x)
  x = x+1

```

---
## For loop

For loops do not have a conditional. Instead you supply an object that you want to be iterate over. This can be a list, tuple, dictionary, set or string. Here we use a list. 

```{python}

x = ['Alpha','Bravo','Charlie']
for i in x:
  print(i)

```


---
## for and range()

The range() function provides us with a nice input for our for loops. It returns a sequence of numbers, starting from 0 and stops before the specified number.

```{python}
for i in range(3):
  print("i is", i)

```

---
## Looping through indices

When we have a numeric range, we can use it to index out from existing objects. This often allows for more complex code evaluation. 

```{python}

geneName = ["Ikzf1","Myc","Igll1"]
expression = [10.4, 4.3, 6.5]
iterations = len(geneName)

for i in range(iterations):
  print(geneName[i]," has an TPM of ",expression[i])

```

---
## Loops and conditionals

Loops can be combined with conditional statements to allow for complex control of their execution over Python objects. 

To help us write complex code we often use pseudocode as a starting point. 


---
## Pseudocode

When we write pseudocode we are trying to write out each computational step in a human readbale way. 

It is important to be specific, simple, concise and include the control structures that would be in your final code. 

```

for 0 to 7
  if value is greater than 5
    print the value and a statement saying it is greater than 5
  else if value is 5
    print the value and a statement saying it is equal to 5
  else if value is less than 5
    print the value and a statement saying it is less than 5


```

---
## Loops and conditionals

Though these can be tough to read and create, by starting with pseudocode and keeping an eye of the hierarchy of indentation we can follow the logic. 

.pull-left[

```
for 0 to 7
  if value is greater than 5
    print the value and a statement saying it is greater than 5
  else if value is 5
    print the value and a statement saying it is equal to 5
  else if value is less than 5
    print the value and a statement saying it is less than 5
```
  ]

.pull-right[

```{python}

for i in range(8):
  if i > 5:
    print("Number",i,"is greater than 5")
  elif i == 5:
    print("Number",i,"is  5") 
  else:
    print("Number",i,"is less than 5") 

```

  ]

---
## Breaking loops

We can use conditionals to exit a loop if a condition is satisfied, just like a while loop.

```{python}
x = range(8)

for i in range(8):
  if i > 5:
    print("Number",i,"is greater than 5")
  elif i == 5:
    print("Number",i,"is  5")
    break
  else:
    print("Number",i,"is less than 5") 

```


---
## Time for an exercise!

??? exercise description here?? [here](../../exercises/exercises/MyExercise2_exercises.html)

---
## Answers to exercise

Answers can be found [here](../../exercises/answers/MyExercise2_answers.html)

---
##  An extra note

Keeping Your Code Nice can be annoying. There exists many ways in which to store your code. Most of the time we are not writing scripts that are production level. Instead you will be doing an analysis of a data set and making decisions in an interactive manner. 

Notebooks give you a means to tie the code, the analysis decisions and the result of the code together into a single file. 

There are several options for python. But Jupyter Notebook is best known for Python. Quarto is growing in popularity as its universal and language agnostic.

--- 
## Further Support

When you hit bugs:
  * Google
  * Stackoverflow
  * Biostars
  * Reach out on [GitHub](https://github.com/RockefellerUniversity/Intro_To_Python/issues)

Other Reference Material:
 * [Harvard's Python Course](https://cs50.harvard.edu/python/2022/)
 * [Geeks For Geeks](https://www.geeksforgeeks.org/getting-started-with-python-programming/)

