---
title: "Introduction to R, Session 2 <html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>"
author: "Rockefeller University, Bioinformatics Resource Centre"
date: "http://rockefelleruniversity.github.io/RU_course_template/"
output: 
  xaringan::moon_reader:
    css: ["default", "metropolisCustom.css", "metropolis-fontsCustom.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
  html_document:
    toc: true # table of content true
    toc_float: yes
    depth: 3  # upto three depths of headings (specified by #, ## and ###)
    number_sections: false  ## if you want number sections at each table header
    theme: united  # many options for theme, this one is my favorite.
    highlight: tango  # specifies the syntax highlighting style
params:
  isSlides: "no"
---
```{r setup, include=FALSE}
suppressPackageStartupMessages(require(knitr))
knitr::opts_chunk$set(echo = TRUE, tidy = T)
```

```{r, results='asis',include=TRUE,echo=FALSE}
if(params$isSlides != "yes"){
  cat("# ?? name of course?? (part 2)

---
"    
  )
  
}

```


##  Recap

?? can add a recap to link to keypoints in first session??
Session 1 covered introduction to R data types and import/export of data.

- [Background to R](https://rockefelleruniversity.github.io/Intro_To_R_1Day/r_course/presentations/singlepage/introToR_Session1.html#background-to-r)
- [Data types in R](https://rockefelleruniversity.github.io/Intro_To_R_1Day/r_course/presentations/singlepage/introToR_Session1.html#data_types_in_r)
- [Reading and writing in R](https://rockefelleruniversity.github.io/Intro_To_R_1Day/r_course/presentations/singlepage/introToR_Session1.html#reading-and-writing-data-in-r)


---
## Same stuff
??More content for second session??

---
## Time for an exercise!


Exercise on functions can be found [here](../..//exercises/exercises/MyExercise3_exercise.html)

---
## Answers to exercise.


Answers can be found here  [here](../..//exercises/answers/MyExercise3_answers.html)

Rcode for answers can be found here  [here](../..//exercises/answers//answers/MyExercise3_Answers.R)


## Slide Structure

We use the a double has/pound for slides. This is important for hierarchy. If you use 1 it will work for the slides, but will mess up the single page version.


## Left/Right alignment

To show things side by side on a slide you can use pull left/right. t is important to keep the tab on the second square bracket. On single page it will just be one after the other:

.pull-left[
This will be on left of slide
  ] 
  
.pull-right[
This will be on right

 ]


## Data from external sources

everything in the rmd is relative to the inst/extdata directory so you have some data to load in, just put it in the data directory, read it in along the path. 
The structure is maintained for those attending the course and download the package using the instructions. The same rules are followed for imgs and scripts that you may call.

```{py}
import 

```

--
## Exercises
The following few slides show you how to structure exercise slides.

We often have several exercise slides per session. So you can just copy and paste and change the directory to the appropriate name. All 3 file types are made from you single exercise Rmd. 

---
## Time for an exercise!

??? exercise description here?? [here](../../exercises/exercises/MyExercise1_exercises.html)

---
## Answers to exercise

Answers can be found [here](../../exercises/answers/MyExercise1_answers.html)

R code for solutions can be found [here](../../exercises/answers/MyExercise1_answers.R)

## Exercise again

So we can have some more sections here describing some new important topic

Then more exercises

---
## Time for an exercise!

??? exercise description here?? [here](../../exercises/exercises/MyExercise2_exercises.html)

---
## Answers to exercise

Answers can be found [here](../../exercises/answers/MyExercise2_answers.html)

R code for solutions can be found [here](../../exercises/answers/MyExercise2_answers.R)