---
title: "My Course Template, Session 1 <html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>"
author: "Rockefeller University, Bioinformatics Resource Centre"
date: "http://rockefelleruniversity.github.io/RU_course_template/"
output: 
  xaringan::moon_reader:
    css: ["default", "metropolisCustom.css", "metropolis-fontsCustom.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
  html_document:
    toc: true # table of content true
    toc_float: yes
    depth: 3  # upto three depths of headings (specified by #, ## and ###)
    number_sections: false  ## if you want number sections at each table header
    theme: united  # many options for theme, this one is my favorite.
    highlight: tango  # specifies the syntax highlighting style
params:
  isSlides: "no"
---

```{r,include=FALSE}
suppressPackageStartupMessages(require(knitr))
knitr::opts_chunk$set(echo = TRUE, tidy = T)
```

```{r, results='asis',include=TRUE,echo=FALSE}
if(params$isSlides != "yes"){
  cat("# ???My Course name (session1)???

---
"    
  )
  
}

```

## Overview

??We often put an overview slide that links to the various parts of the course content to make it easy to navigate i.e. this one for intro to r??

- [Set up](https://rockefelleruniversity.github.io/Intro_To_R_1Day/r_course/presentations/singlepage/introToR_Session1.html#set-up)
- [Background to R](https://rockefelleruniversity.github.io/Intro_To_R_1Day/r_course/presentations/singlepage/introToR_Session1.html#background-to-r)
- [Data types in R](https://rockefelleruniversity.github.io/Intro_To_R_1Day/r_course/presentations/singlepage/introToR_Session1.html#data_types_in_r)
- [Reading and writing in R](https://rockefelleruniversity.github.io/Intro_To_R_1Day/r_course/presentations/singlepage/introToR_Session1.html#reading-and-writing-data-in-r)

---
```{r, results='asis',include=TRUE,echo=FALSE}
if(params$isSlides == "yes"){
  cat("class: inverse, center, middle

# Set Up

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=720px></html> 

---
"    
  )
}else{
  cat("# Set Up

---
"    
  )
  
}

```

## Materials

??We try to denote each major section of content with these title slides (see above for set up). The hierarchy will help with the different versions of the teaching content i.e. slides and single page.??

?? Materials just gives the links to download the package Just ensure links are up to date??

All prerequisites, links to material and slides for this course can be found on github.

* [Intro_To_R_1](https://rockefelleruniversity.github.io/Intro_To_R_1Day/)

Or can be downloaded as a zip archive from here.

* [Download zip](https://github.com/rockefelleruniversity/Intro_To_R_1Day/zipball/master)

---
## Course materials

?? this slide just shows where thye can get the course content from the download. You do not need to populate these directories. It will be made during compilation??

Once the zip file in unarchived. All presentations as HTML slides and pages, their R code and HTML practical sheets will be available in the directories underneath.

* **r_course/presentations/slides/**
Presentations as an HTML slide show.
* **r_course/presentations/singlepage/** 
Presentations as an HTML single page.
* **r_course/presentations/r_code/**
R code in presentations.
* **r_course/exercises/**
Practicals as HTML pages. 
* **r_course/answers/**
Practicals with answers as HTML pages and R code solutions. 


---
## Set the Working directory


Before running any of the code in the practicals or slides we need to set the working directory to the folder we unarchived. 

You may navigate to the unarchived RU_Course_help folder in the Rstudio menu.

**Session -> Set Working Directory -> Choose Directory**

or in the console.

```{r setwd_introtoR,eval=F} 
setwd("/PathToMyDownload/RU_Course_template/r_course")
# e.g. setwd("~/Downloads/Intro_To_R_1Day/r_course")
```


---
```{r, results='asis',include=TRUE,echo=FALSE}
if(params$isSlides == "yes"){
  cat("class: inverse, center, middle

# Another section header

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=720px></html> 

---
"    
  )
}else{
  cat("# Another section header

---
"    
  )
  
}

```

## Slide Structure

We use the a double has/pound for slides. This is important for hierarchy. If you use 1 it will work for the slides, but will mess up the single page version.


## Left/Right alignment

To show things side by side on a slide you can use pull left/right. t is important to keep the tab on the second square bracket. On single page it will just be one after the other:

.pull-left[
This will be on left of slide
  ] 
  
.pull-right[
This will be on right

 ]


## Data from external sources

everything in the rmd is relative to the inst/extdata directory so you have some data to load in, just put it in the data directory, read it in along the path. 
The structure is maintained for those attending the course and download the package using the instructions. The same rules are followed for imgs and scripts that you may call.
```{r,echo=T}
Table <- read.table("data/readThisTable.csv",sep=",",header=T)

```

## Big Processes

If what you want to do is super intense computationally you can use the eval=F arguments to prevent it being done during compilation.

```{r,echo=T, eval=F}
#Intense computation
myresult<-10^6 +1

```

You can then just load in the result of the computation, and keep using the result after that. If you do want them to evaluate the intense computation, then you can hide this intermediate file that you are loading in with echo=F. 

```{r,echo=F, eval=T}
load("data/myresult.RData")

```

--
## Exercises
The following few slides show you how to structure exercise slides.

We often have several exercise slides per session. So you can just copy and paste and change the directory to the appropriate name. All 3 file types are made from you single exercise Rmd. 

---
## Time for an exercise!

??? exercise description here?? [here](../../exercises/exercises/MyExercise1_exercises.html)

---
## Answers to exercise

Answers can be found [here](../../exercises/answers/MyExercise1_answers.html)

R code for solutions can be found [here](../../exercises/answers/MyExercise1_answers.R)

## Exercise again

So we can have some more sections here describing some new important topic

Then more exercises

---
## Time for an exercise!

??? exercise description here?? [here](../../exercises/exercises/MyExercise2_exercises.html)

---
## Answers to exercise

Answers can be found [here](../../exercises/answers/MyExercise2_answers.html)

R code for solutions can be found [here](../../exercises/answers/MyExercise2_answers.R)
