---
title: "Exercise 4"
author: "Rockefeller University, Bioinformatics Resource Centre"
date: "https://rockefelleruniversity.github.io/Intro_To_Python/"
output: 
  html_document:
    number_sections: false  ## if you want number sections at each table header
    theme: united  # many options for theme, this one is my favorite.
    highlight: tango  # specifies the syntax highlighting style
params:
  toMessage: false    
---

```{r, setup,include=FALSE}
toMessage <- ifelse(!is.null(params$toMessage),params$toMessage,FALSE)
message(toMessage)
```


<p>&nbsp;</p>
These exercises are about logical operators and reading/writing from [Session 2](https://rockefelleruniversity.github.io/Intro_To_Python/). 

**Exercise 1** - **If Else**

- Build an if statement that test if x is a negative number. Write a print statement that will confirm if x is less than 0. Test it with several values of x.


```{python,echo=toMessage}
x=42

if x<0:
  print("It's a negative number!")

x=-42

if x<0:
  print("It's a negative number!")

```

- Modify the previous if statement to include an else. Ensure a print statement is returned to say if the number is not negative. 

```{python,echo=toMessage}

x=0

if x<0:
  print("It's a negative number!")
else:
  print("It's not a negative number!")

x=-1

if x<0:
  print("It's a negative number!")
else:
  print("It's not a negative number!")

```

- Finally add a elif statement. We want a response to confirm if x is negative, positive or if it is zero. 

```{python,echo=toMessage}
x=-1

if x<0:
  print("It's a negative number!")
elif x==0:
  print("It's zero")
else:
  print("It's a positive number!")
  
x=0

if x<0:
  print("It's a negative number!")
elif x==0:
  print("It's zero")
else:
  print("It's a positive number!")
  
x=1

if x<0:
  print("It's a negative number!")
elif x==0:
  print("It's zero")
else:
  print("It's a positive number!")

```

- Build an if/else statement that test if a variable is odd/even. Include x in the printed output. 

*Hint1: x will likely be an integer so it will need to be coerced to print out as part of a message string.*
*Hint2: The modulus operator may be useful here i.e. x%2 returns the remainder after the value of x is divided by 2.* 


```{python,echo=toMessage}
x=1

if x%2==0:
  mymessage = str(x) + " is even"
  print(mymessage)
else:
  mymessage = str(x) + " is odd"
  print(mymessage)
  
  
x=42

if x%2==0:
  mymessage = str(x) + " is even"
  print(mymessage)
else:
  mymessage = str(x) + " is odd"
  print(mymessage)

```

- Put this in a function and try out it out on various variables.

```{python,echo=toMessage}

def even_or_odd(x):
  if x%2==0:
    mymessage = str(x) + " is even"
    print(mymessage)
  else:
    mymessage = str(x) + " is odd"
    print(mymessage)

even_or_odd(10)
even_or_odd(3)
even_or_odd(10007)
```

**Exercise 2** - **Loops**

- Calculate the factorial (factorial of 3 = 3 * 2 * 1) of 10 using a loop. 
*Hint: first try to write some pseudocode to help you build this*

```{r, echo=toMessage, eval=F}

# Idea
Each iteration of a loop I want to multiply my factorial by itself:
x=1  result_1 = 1
x=2  result_2 = result_1*x = 2*1
x=3  result_3 = result_2*x = 3*2*1
x=4  result_4 = result_3*x = 4*3*2*1
etc...

# Pseudocode

for 0 to 9:
  if x is 0:
    create an object and set its value to 1
  else:
    once we are past the first data point we can multiply factorialAnswer by the iteration we are on. Then overwrite factorialAnswer with this result. 

```

```{python,echo=toMessage}

for x in range(9):
  if x == 0:
    factorialAnswer = 1
  else:
    factorialAnswer = factorialAnswer * (x+1)

factorialAnswer
```

- Adjusting your answer from before, what is the first number that has a factorial greater than 1000?
*Hint: A while loop will be good for this*

```{python,echo=toMessage}

factorialAnswer = 0
count = 0

while factorialAnswer <= 1000:
  if count == 0:
    factorialAnswer = 1
  else:
    factorialAnswer = factorialAnswer * (count+1)
  count = count+1
    
    

count
factorialAnswer

```

